\section{Usage and Implementation}

Our compiler supports two ways to use the reference file feature, with runtime calls or with compiler directives.
The directives simplify the usage quite a bit, since the compiler knows things like the datatype, file and function names, and line number, and can pass this information to the compare routines automatically.
In either case, the user introduces runtime calls or compiler directives at the points in the program where intermediate or final results should be compared.
The directives must be enabled with a command line flag, and will be ignored by other compilers.
This allows a user to leave the directives in place, making their insertion part of the continuing maintenance procedure.
The runtime calls can be placed in conditionally compiled regions, or will have to be isolated in a testing harness, or removed before releasing the software.

The directives use Fortran array notation or the same C array section notation used in OpenACC and OpenMP.
In Fortran, the user can insert a directive like:
\begin{otabbing}
\>\textbt{!$pgi compare(a)}
\end{otabbing}
where \textbt{a} is a Fortran simple variable or array.
The user can have two or more names on a single directive, and can use array sections as well:
\begin{otabbing}
\>\textbt{!$pgi compare(a(:),b(1:100,2:20))}
\end{otabbing}
In C++ or C, the user can insert a pragma like:
\begin{otabbing}
\>\textct{pragma pgi compare(x,a[0:n])}
\end{otabbing}
If \textbt{x} is a scalar, a fixed-size array or C VLA, the compiler knows the size of the array; otherwise, the user must use the array section notation to tell the compiler the size of a dynamically allocated array.
The compiler will interpret the directives, essentially replacing them with runtime calls that handle each array, implicitly including the filename, function name, line number, variable or array name, and datatype information, for more precise error reporting.
There are additional clauses that can appear on the directive, to specify the tolerance to use and how to proceed in case of an error.
The directives are enabled with the \textbt{-Mcompare} compiler flag.

The alternative is to use runtime calls.
In Fortran, the user can insert
\begin{otabbing}
\>\textbt{call pgi_compare(a, n, 'real')} \\
\>\textbt{call pgi_compare(b, 20000, 'double precision')}
\end{otabbing}
In C++ or C, the user would insert the function call:
\begin{otabbing}
\>\textbt{pgi_compare(x, 1, "float",__FILE__,__LINE__)} \\
\>\textbt{pgi_compare(a, n, "double",__FILE__,__LINE__)}
\end{otabbing}
This method requires the user to specify the datatype in a string, and won't get the other identifying information automatically, but the compare procedure is otherwise identical.
This method also loses the ability to specify different tolerances and behavior for different arrays.

Whether using directives or runtime calls, the program will save the values to a reference file, or compare the values to those saved in that file.
In the test run, the runtime routine will validate the program sequence by checking that any available filename and line number information matches, as well as that the saved data sequence matches in datatype and number of values.
The compare operation is controlled by the directive clauses and by the \textbt{PGI_COMPARE} environment variable.
\textbt{PGI_COMPARE} must have a sequence of comma-separated options,
shown in the table in Table~\ref{env}.
By default, if the file does not exist, it is created;
if it does exist, it is treated as the reference file against which to compare.
\begin{table}
\begin{center}
\begin{tabular}{ll}
\hline
option & Description \\
% \multicolumn{1}{l}{\rule{0pt}{12pt} \textbt{export PGICOMPARE=option[,option]}
% }&\multicolumn{1}{l}{  Description }\\[2pt]
\hline
\textbt{FILE=}\textit{filename} & Name of reference file \\
\textbt{CREATE}   &   This run creates the reference file \\
\textbt{COMPARE}   &   Compare this run to the reference file \\
\textbt{ABS=}\textit{r} & Use \textit{r} as an absolute tolerance \\
\textbt{REL=}\textit{r} & Use \textit{r} as a relative tolerance \\
\textbt{ULP=}\textit{n} & Allow \textit{n} differences in ULP \\
\textbt{IEEE} & Test IEEE NaNs\\
\textbt{REPORT=}\textit{n} & Report first \textit{n} differences \\
\textbt{SKIP=}\textit{n}    & Skip the first \textit{n} differences \\
\textbt{VERBOSE}   & Outputs all details of comparison \\
\textbt{PATCH}   &   Patch erroneous values with correct values \\
\textbt{STOP}   &   Stop at first difference \\
\hline
\end{tabular}
\end{center}
\caption{Options that can appear in the \textbt{PGI_COMPARE} environment variable.}
\label{env}
\end{table}

When comparing integer values, the tolerance is ignored and the test value must match the reference value exactly.
When comparing floating point values, the test value will be compared using an absolute tolerance, a relative tolerance, compare ignoring the last ULP digits, or any combination of the three.
If no tolerance is specified, the tolerance is zero.
If any tolerance is exceeded, then the comparison fails.
If the test value is a \emph{NaN} but does not match the reference value, the comparison fails as well.
This comparison scheme was copied from what we use in our nightly and weekly quality assurance tests, where we run over 1,000,000 tests every weekend, plus more during shorter nightly runs.
If a comparison fails, the program can ignore some number of errors or report on them, then either stop, continue execution, or patch the bad errors and then continue.
